Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States

### 2017-11-29 doc/manuscript/manuscript.Rmd
add Sydney's comments

All data processing and analyses were performed in a Rocker "verse" project container running `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@boettiger2017introduction; @R]. 
These analyses were rendered as dynamic documents with the R packages *knitr* (version `r packageVersion("knitr")`) and *ezknitr* (version `r packageVersion("ezknitr")`) and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/ [@attali2016ezknitr; @xie2017knitr].
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
Loci with compound repeats were removed for the reported statistics.
To ensure the integrity of the results we additionally processed these loci and included them in concurrent analyses.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotia from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that duplicated genotypes were reduced to a single observation when they were collected in the same year from the same host cultivar located in the same source field (wmn or producer)---for subsequent analysis. 
We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], evenness ($E_5$), and the expected number of multilocus haplotypes (*eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both haplotype diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].

@@

The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present. 
Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (See section on 'Loading Data and Setting Strata' in the MLG-distribution.md[^1] file in the supplemental files [@kamvar2017data]), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.

@@

loc_table %>%
  mutate(allele = round(allele, 2)) %>%
  mutate(Repeats = case_when(
    Locus == "5-2"   ~ "(GT)",
    Locus == "6-2"   ~ "(TTTTTC)(TTTTTG)(TTTTTC)", # "(TTTTTC)~2~(TTTTTG)~2~(TTTTTC)",
    Locus == "7-2"   ~ "(GA)",
    Locus == "8-3"   ~ "(CA)",
    Locus == "9-2"   ~ "(CA)(CT)", # "(CA)~9~(CT)~9~",
    Locus == "12-2"  ~ "(CA)",
    Locus == "17-3"  ~ "(TTA)",
    Locus == "20-3"  ~ "(GT)GG(GT)", # "(GT)~7~GG(GT)~5~",
    Locus == "55-4"  ~ "(TACA)",
    Locus == "110-4" ~ "(TATG)",
    Locus == "114-4" ~ "(TAGA)"
  )) %>%
  rename(`*h*` = Hexp) %>%
  rename(`No. alleles` = allele) %>%
  rename(`Repeat Motif` = Repeats) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
  huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
  huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
  huxtable::print_md(max_width = 100)
```

Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation]. Average *h* = `r signif(Hmean, 3)`, average Evenness = `r signif(Emean, 3)`, average no. alleles = `r signif(Amean, 3)`

>>>>

All data processing and analyses were performed in a Rocker "verse" project container running `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@boettiger2017introduction; @R]. 
These analyses were rendered as dynamic documents with the R packages *knitr* (version `r packageVersion("knitr")`) and *ezknitr* (version `r packageVersion("ezknitr")`) and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/ [@attali2016ezknitr; @xie2017knitr].
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
Loci with compound repeats were removed for the reported statistics.
To ensure the integrity of the results we additionally processed these loci and included them in concurrent analyses.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
We additionally assessed within-locus allelic diversity by measuring Nei's gene diversity (*h*) [@nei1978estimation] and allelic evenness ($E_5$) [@pielou1975ecological; @grunwald2003analysis].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotia from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that duplicated genotypes were reduced to a single observation when they were collected in the same year from the same host cultivar located in the same source field (wmn or producer)---for subsequent analysis. 
We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], evenness ($E_5$), and the expected number of multilocus haplotypes via rarefaction (*eMLH*) with 10 samples [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
If all haplotypes in are equally abundant, both *G* and $e^H$ (the exponentiation of *H*) are expected to be equal to the number of haplotypes, $\lambda$ and $E_5$ are expected to equal one, and *eMLH* is expected to be at its maximum value (in this case, 10) [@grunwald2003analysis].
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both haplotype diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].

@@

The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and $E_5$ (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present. 
Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (See section on 'Loading Data and Setting Strata' in the MLG-distribution.md[^1] file in the supplemental files [@kamvar2017data]), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.

@@

loc_table %>%
mutate(allele = round(allele, 2)) %>%
mutate(Repeats = case_when(
Locus == "5-2"   ~ "(GT)",
Locus == "6-2"   ~ "(TTTTTC)(TTTTTG)(TTTTTC)", # "(TTTTTC)~2~(TTTTTG)~2~(TTTTTC)",
Locus == "7-2"   ~ "(GA)",
Locus == "8-3"   ~ "(CA)",
Locus == "9-2"   ~ "(CA)(CT)", # "(CA)~9~(CT)~9~",
Locus == "12-2"  ~ "(CA)",
Locus == "17-3"  ~ "(TTA)",
Locus == "20-3"  ~ "(GT)GG(GT)", # "(GT)~7~GG(GT)~5~",
Locus == "55-4"  ~ "(TACA)",
Locus == "110-4" ~ "(TATG)",
Locus == "114-4" ~ "(TAGA)"
)) %>%
rename(`*h*` = Hexp) %>%
rename(`$E_5$` = Evenness) %>%
rename(`No. alleles` = allele) %>%
rename(`Repeat Motif` = Repeats) %>%
huxtable::as_huxtable(add_colnames = TRUE) %>%
huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
huxtable::print_md(max_width = 100)
```

Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's 1978 gene diversity, $E_5$ = Evenness. Average *h* = `r signif(Hmean, 3)`, average $E_5$ = `r signif(Emean, 3)`, average no. alleles = `r signif(Amean, 3)`
